### 2017-05-30 00-Abstract.Rmd
init

# Résumé et Abstract {-}

Écrire le résumé français ici...

Write the english abstract here...