### 2017-05-30 02-Introduction.Rmd

# Introduction {-}

* Introduce why tropical rainforest are such a good study case
* Define biodiversity 
* Define ecosystem functionning
* Define ecosystem services
* Introduce tropical sylviculture
* Define selective logging
* Introduce biodiversity erosion in tropical rainforests ant its impact
* Introduce Loreau partitionning
* Introduce the interest of modelling
* Introduce the idea to look at biodiversity as both a parameter and an output

Tropical rainforests have fascinated ecologists due to their outstanding diversity [@connell_diversity_1978].

Tropical forests are primary ecosystems in terms of biodiversit and carbon storage [@Lewis2004].

High biodiversity from tropical rainforests is the source of many ecosystem functions, which support ecosystem services.

Biodiversity and ecosystem functionning field is emmering [@Loreau2000].

Selective logging is increasing in tropical forests.

Tropical logging accounts for one eight of global timber production [@Blaser2011].

Selective logging represents a major threat to biodiversity [@Gibson2013].