### 2017-05-30 03-MnMs.Rmd

# Materials and methods

## TROLL simulator

### Abiotic environment

### Photosynthesis

### Autotrpohic respiration

### Carbon uptake

### Tree growth

### Seed dispersion, production and recruitment

### Mortality

## TROLL sensitivity analysis

### Functional traits

### Seed rain

## Disturbance

### Disturbance module

### Design of experiment

### Outputs anlaysis ?

#### Resistance and resilience metrics

#### Biodiversity partitioning

## Selective logging

### Selective logging module

#### Designation

#### Selection

#### Rotten trees

#### Felling

#### Tracks

#### Gap damages

### Design of experiment

### Outputs analysis ?

#### Resistance and resilience metrics

#### Biodiversity partitioning