### 2017-05-30 latex/before_body.tex

\date{}

\thispagestyle{empty}
\begin{center}
\includegraphics{images/main.pdf}
\end{center}

\setlength{\abovedisplayskip}{-5pt}
\setlength{\abovedisplayshortskip}{-5pt}