### 2017-05-30 03-Model.Rmd

# Model description

## Overview

## Abiotic environment

## Photosynthesis

## Autotrpohic respiration

## Carbon uptake

## Tree growth

## Seed dispersion, production and recruitment

## Mortality