### 2017-05-30 index.Rmd

--- 
site: "bookdown::bookdown_site"
title: "Master thesis"
author: "Sylvain SCHMITT"
date: "`r Sys.Date()`"
output: bookdown::gitbook
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
colorlinks: yes
cover-image: images/main.png
---

`r if (knitr:::is_html_output()) '# Preface {-}'`

`r if (knitr:::is_html_output()) '![stability-wip](https://img.shields.io/badge/stability-work_in_progress-lightgrey.svg)'`

```{r fig.align='center', echo=FALSE, include=identical(knitr:::pandoc_to(), 'html')}
knitr::include_graphics('images/main.png', dpi = NA)
```

`r if (knitr:::is_html_output()) 'This is the online html version of my master thesis. The pdf version is available for download with the download button on top of the windows.'`

`r if (knitr:::is_html_output()) 'Have a nice reading.'`

`r if (knitr:::is_html_output()) '<div align="right">  *Sylvain*'`