### 2017-06-02 latex/after_body.tex
template tex v2

%Last page
\newpage
\paragraph{Résumé :}
Écrire le résumé ici...
\paragraph{Mots clés :} mots clés
\newline\newline
\paragraph{Abstract:}
Write abstract here
\paragraph{Keywords:} keywords

\vspace*{\fill}
\includegraphics{images/logo}