### 2017-06-02 01-Acknowledgments.Rmd
Model and M&M reorganized + sylviculture model v1

# Acknowledgments {-}

I would like to thank...

- Bruno
- Stéphane
- Laurent
- Fabian
- Isabelle
- Jérôme
- Camilla
- Aurélie
- Éric
- ...