### 2017-06-09 09-Appendix3.Rmd

# Appendix 3: Rotten tree model

In order to simulate sylviculture with TROLL we need to implement a new sylviculture module inside TROLL model code. A first litterature review was completed by an interview with Laurent Descroix of the Office Nationale des Forêts. We discovered that rotten trees were not random and seemed to depend both on tree species and diameter. This document presents modelling of relation between rotten trees and their species and diameter.

In fact we have two different questions:

- Predict if a tree will be probed as rotten (models **M**) 
- Predict how much of tree volume is rotten (models **N**)

First all **M** model can be written as follow:
$$Rotten_n \sim \mathcal{B}(\theta_n), ~~n \in [1,N_{=3816}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]$$
Secondly, all **N** models depend on a latent variable being the percentage of rotten wood $Pt_r$. We can assume that all trees are growing depending on species $s$ and plot $p$ fertility and are supposed to have a full healthy volume $V_h$ for a given diameter $dbh$. We obtain following model:

$$V_f \sim \mathcal{logN}(V_h*Pt_r, \sigma), ~~n \in [1,N_{=3268}] ~~ p \in [1, P_{=8}], ~~ s \in [1, S_{=43}]$$

We retained following models :

```{r A3models, echo=FALSE}
mpath <- "~/Documents/ECOFOG/TROLL/inst/doc/rotten_models"
mtab <- data.frame(
  M = c('$M_{s,p}$', '$N_{s,p} + L_{s,p}$'),
  Model = c(
    "${P_{rotten}}_n \\sim \\mathcal{B}(inv_{logit}(\\beta_0 + \\beta_1*dbh_n + {\\beta_2}_p + {\\beta_3}_s))$",
    "$Volume_{of~wood} \\sim \\mathcal{logN} (log[((\\beta + \\beta_p + \\beta_s)*dbh^2)*(1 - Pr*((\\theta + \\theta_p + \\theta_s) *dbh^2))], \\sigma)$"
    )
  )
mtab <- apply(mtab, 2, as.character)
knitr::kable(mtab, caption = 'Models summary.', format = 'pandoc')
row.names(mtab) <- c('Mps', 'Nps')
```

## Probed rotten (M)

Based on complexity (number of parameters), convergence and likelihood we selected model $M_{p,s}$:

`r paste(mtab['Mps',], collapse = ': ')`

```{r A3M, include=FALSE, cache=TRUE}
library(dplyr)
data <- read_csv("~/Documents/ECOFOG/TROLL/inst/extdata/Rotten/rotten_data.csv")
data <- data[c('num', 'Parcelle', 'up', 'ess', 'abattu', 'tout BE', 'sonde_creux', 'sonde_creux abatt', 'sonde_creux_refus', 'DBH_desi', 'V_Est', 'vol_BO')]
names(data) <- c('id', 'plot', 'subplot', 'spCode', 'logged', 'rotten_full', 'probed_rotten', 'probed_rotten_logged', 'probed_rotten_refused', 'dbhest', 'Vest', 'Vbo')
species <- read_csv("~/Documents/ECOFOG/TROLL/inst/extdata/Rotten/rotten_species.csv")[-1,]
species <- species[c("Code \nEssence" , "Nom \nscientifique")]
names(species) <- c('spCode', 'species')
species$wsg <- apply(do.call('rbind', lapply(strsplit(species$species, ' ', fixed = T), function(x) x[1:2])),
                   1, function(x) getWoodDensity(x[1], x[2], region = 'SouthAmericaTrop')$meanWD)
data <- data.frame(left_join(data, species)) ; rm(species)
data <- data[-which(is.nan(data$wsg)),]
data$sp <- apply(data[c('plot', 'species')], 1 , paste, collapse = '.')
dbhmax <- max(data$dbhest)
data$dbh <- data$dbhest/dbhmax
wsgmax <- max(data$wsg)
load(file.path(mpath, 'Mps.Rdata'))
data2 <- group_by(data, dbhest) %>%
  summarise_each(funs(mean))
inv_logit <- function(u) {
  return(1 / (1 + exp(-u)));
}
beta <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),c('beta_0', 'beta_1')]
Pr <- function(dbh, dbhmax, beta0, beta1){
  theta <- beta0 + beta1*(dbh/dbhmax)
  inv_logit(theta)
}
pred <- apply(as.matrix(fit), 1, function(x)
  sapply(seq_len(dim(data2)[1]), function(i)
    Pr(data2$dbhest[i], dbhmax, x['beta_0'], x['beta_1'])
  ))
pred <- data.frame(t(apply(pred, 1, function(x) quantile(x, probs = c(0.05, 0.95)))))
names(pred) <- c('5%', '95%')
pred$mean <- inv_logit(beta[1] + (beta[2]/dbhmax)*data2$dbhest)
pred$dbh <- data2$dbhest
```

```{r A3Mpred, echo=FALSE, caption='Models prediction. Probability to be probed as rotten P for a given diameter at breast hieght (cm).'}
library(ggplot2)
ggplot(data2, aes(x = dbhest, y = probed_rotten)) +
  geom_ribbon(aes(x = pred$dbh, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
  geom_line(aes(x = pred$dbh, y = pred$`5%`, color = '5%'), linetype = 2) +
  geom_line(aes(x = pred$dbh, y = pred$`95%`, color = '95%'), linetype = 2) +
  geom_line(aes(x = pred$dbh, y = pred$mean, color = 'mean'), linetype = 2) +
  geom_point() +
  theme_bw() +
  xlab('Diameter at breast height (cm)') +
  ylab('Probability to be probed as rotten') +
  scale_color_manual(name = 'Model',
                     values = c('5%' = 'black',
                                '95%' = 'black',
                                'mean' = 'red'))
```

```{r A3Mtab, echo=FALSE}
table <- data.frame(row.names = seq(45, 125, 5),
           `P` = round(sapply(seq(45, 125, 5), function(dbh) 100*inv_logit(beta[1]+beta[2]*dbh/dbhmax))))
knitr::kable(t(table), caption = 'Models prediction. Probability to be probed as rotten (P in %) for a given dbh (cm).', format = 'pandoc')
```

## Rotten volume (N)

Based on complexity (number of parameters), convergence and likelihood we selected model $N_{p,s}$ associated to hyperparameter $\rho$ with model $L_{p,s}$:

`r paste(mtab['Nps',], collapse = ': ')`

```{r A3N, include=FALSE, cache=TRUE}
data <- read_csv("~/Documents/ECOFOG/TROLL/inst/extdata/Rotten/rotten_data.csv")
data <- data[c('num', 'Parcelle', 'up', 'ess', 'abattu', 'tout BE', 'sonde_creux', 'sonde_creux abatt', 'sonde_creux_refus', 'DBH_desi', 'V_Est', 'vol_BO')]
names(data) <- c('id', 'plot', 'subplot', 'spCode', 'logged', 'rotten_full', 'probed_rotten', 'probed_rotten_logged', 'probed_rotten_refused', 'dbhest', 'Vest', 'Vbo')
data <- subset(data, !is.na(Vbo))
data <- subset(data, Vbo!=0)
data <- data[c('id', 'plot', 'spCode', 'probed_rotten', 'Vbo', 'dbhest')]
species <- read_csv("~/Documents/ECOFOG/TROLL/inst/extdata/Rotten/rotten_species.csv")[-1,]
species <- species[c("Code \nEssence" , "Nom \nscientifique")]
names(species) <- c('spCode', 'species')
data <- data.frame(left_join(data, species)) ; rm(species)
dbhmax <- max(data$dbhest)
data2 <- data[sample.int(dim(data)[1], 500, replace = F),]
load(file.path(mpath, 'Lps.Rdata'))
pars <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),c('beta', 'sigma', 'theta')]
Vf <- function(dbh, dbhmax, Pr, pars){
  rho <- pars[3]*(dbh/dbhmax)*(dbh/dbhmax)
  mu <- (pars[1]*(dbh/dbhmax)*(dbh/dbhmax))*(1-Pr*rho)
  rlnorm(1, log(mu), pars[2])
}
pred <- apply(as.matrix(fit), 1, function(x)
  sapply(seq_len(dim(data2)[1]), function(i)
    Vf(data2$dbhest[i], dbhmax, data2$probed_rotten[i], c(x['beta'], x['sigma'], x['theta']))
  ))
pred <- data.frame(t(apply(pred, 1, function(x) quantile(x, probs = c(0.05, 0.95), na.rm = T))))
names(pred) <- c('5%', '95%')
pred$mean <- (pars[1]*(data2$dbhest/dbhmax)*(data2$dbhest/dbhmax))*(1-data2$probed_rotten*(pars[3]*(data2$dbhest/dbhmax)*(data2$dbhest/dbhmax)))
pred$dbh <- data2$dbhest
```

```{r A3Npred, echo=FALSE, caption='Models prediction. Final volume of wood for a given diameter at breast hieght (cm).'}
ggplot(data2, aes(x = dbhest, y = Vbo)) +
  geom_ribbon(aes(x = pred$dbh, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
  geom_line(aes(x = pred$dbh, y = pred$`5%`, color = '5%'), linetype = 2) +
  geom_line(aes(x = pred$dbh, y = pred$`95%`, color = '95%'), linetype = 2) +
  geom_line(aes(x = pred$dbh, y = pred$mean, color = 'mean'), linetype = 2) +
  geom_point() + geom_jitter() +
  theme_bw() +
  xlab('Estimated dbh (cm)') +
  ylab('Final softwood product (m3)') +
  scale_color_manual(name = 'Model',
                     values = c('5%' = 'black',
                                '95%' = 'black',
                                'mean' = 'red'))
```

```{r A3Ntab, echo=FALSE}
table <- data.frame(row.names = seq(45, 125, 5),
           Vf = round(sapply(seq(45, 125, 5), function(dbh) (pars[1]*(dbh/dbhmax)^2)*(1-pars[3]*(dbh/dbhmax)^2)),2),
           Vp = round(sapply(seq(45, 125, 5), function(dbh) 100*pars[3]*(dbh/dbhmax)^2)))
knitr::kable(t(table), caption = 'Models prediction. Final volume of wood ($V_f$ in $m^3$) and percent of rotten wood ($V_p$ in %) for a given dbh (cm) if the tree was probed rotten.', format = 'pandoc')
```