### 2017-07-13 01-Acknowledgments.Rmd
Thesis v1

# Acknowledgments {-}

I would like to thank...

- Bruno
- Stéphane
- Laurent
- Fabian
- Isabelle
- Jérôme
- Camilla
- Aurélie
- Éric
- Maxime
- Raphaël
- Paracou team
- ...