### 2017-07-19 01-Acknowledgments.Rmd

# Acknowledgments {-}

I would like to thank...

- Bruno
- Stéphane
- Camilla
- Fabian
- Isabelle
- Laurent
- Jérôme
- Aurélie
- Éric
- Maxime
- Raphaël
- AMAP colleagues
- Myriam
- BIOGECO colleagues
- Pascal Pe
- Paracou team
- Pascal Pa
- Aurélie C