### 2018-11-06 SX-mortality.Rmd

---
title: "Nino mortality issue"
date: '`r Sys.Date()`'
output:
  bookdown::pdf_document2:
    number_sections: false
    toc: false
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
theme_set(bayesplot::theme_default())
opts_chunk$set( echo = T, message = F, warning = F, 
                fig.height = 8, fig.width = 12,
                cache = T, cache.lazy = F)
```

This document explore if the issue encoutered by Nino of over-mortality in years following the selective logging appeared in my simulations too in TROLL version 2.3.1 with my previous version of the sylviculture module.

```{r data}
path <- '~/Documents/ECOFOG/Results/disturbance/sylviculture/data/'
load(file.path(path, 'ecosystem.Rdata'))
ecosystem <- ecosystem %>% 
  separate(sim, c("SR", "CHV", "control"), remove = F, extra = "drop") %>% 
  filter(SR == 125, is.na(control))
Vkilled <- ecosystem %>% 
  group_by(sim) %>% 
  filter(time %in% 1:2) %>% 
  arrange(time) %>% 
  summarise(Vkilled = 1-last(agb)/first(agb)) %>% 
  select(Vkilled) %>% unlist()
```

```{r plot}
ggplot(ecosystem, aes(time, agb, color = sim)) +
  geom_point() +
  geom_line() +
  ggtitle(paste("Volume killed:", round(mean(Vkilled*100), 2), 
            "%", "(+/-", round(sd(Vkilled*100), 2), "%)")) +
  scale_color_discrete(guide = F) +
  xlim(0,10)
```