### 2019-01-03 A02-Resilience.Rmd
Draft 0.1.9011 JC and CP review except discussion last paragraph

---
title: "A02 - Resilience index"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::pdf_document2:
    number_sections: false
    toc: false
  bookdown::word_document2: default
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
theme_set(bayesplot::theme_default())
opts_chunk$set( echo = F, message = F, warning = F, 
                fig.height = 8, fig.width = 8,
                cache = T, cache.lazy = F)
```

```{r resilience, echo=FALSE}
rich <- c(5,25,125)
cvh <- 1:20
dist <- c(0,25,50,75)
path <- '~/Documents/ECOFOG/Results/disturbance/disturbance/data'
load(file.path(path, 'disturbanceDOE.Rdata'))
doe <- rename(doe, SR = richness)
load(file.path(path, 'disturbanceData.Rdata'))
load(file.path(path, 'disturbanceResilience.Rdata'))
load(file.path('~/Documents/ECOFOG/Results/disturbance/maturity', 'matureData.Rdata'))
mature <- datal$time %>% mutate(time = time - 600) ; rm(datal)
```

```{r Deq, cache=TRUE, echo=FALSE}
Deq <- resilience %>% 
  group_by(time, sim) %>% 
  mutate_all(funs((1-.)^2)) %>% 
  mutate(structure = sqrt(sum(agb, ba, n, n10, n30))) %>% 
  mutate(production = sqrt(sum(gpp, npp, Rday, Rnight)))
```

```{r Ir, cache=TRUE, echo=FALSE}
Ir <- Deq %>% 
  group_by(sim) %>% 
  arrange(time) %>% 
  mutate_at(vars(-time, -sim), cumsum) %>% 
  filter(time == 600)
```

```{r IrGraph1, eval=F, include=F}
mature %>% 
  select(time, sim, agb) %>% 
  separate(sim, c("SR", "CVH"), remove = F) %>% 
  mutate(disturbance = 50) %>% 
  rbind(ecosystem %>% 
          select(time, sim, agb) %>% 
          separate(sim, c("SR", "CVH", "disturbance"), remove = F)) %>% 
  filter(SR == 25, CVH == 4, disturbance %in% c(0, 50)) %>% 
  mutate(agb = agb/10^3) %>% 
  reshape2::dcast(time ~ disturbance, value.var = "agb") %>% 
  rename(control = `0`, treatment = `50`) %>% 
  ggplot(aes(x = time)) +
  geom_line(aes(y = control, color = "control")) +
  geom_line(aes(y = treatment, color = "disturbed")) +
  geom_ribbon(aes(ymax = control, ymin = treatment), 
              fill = "lightgrey", col = NA, alpha = 0.5) +
  xlab("Time (years)") +
  ylab("Aboveground biomass (tonC/ha)") +
  scale_color_discrete("Disturbance\nintensity\n(% of BA)", 
                       labels = c("0 (control)", "50 (disturbed)")) #+
  # ggsave(filename = "./images/original1.png", width = 20, height = 10, units = "cm")
```

```{r IrGraph2, eval=F, include=F}
bind_rows(
  mature %>% 
    select(time, sim, agb) %>% 
    separate(sim, c("SR", "CVH"), remove = F) %>% 
    mutate(disturbance = 50) %>% 
    rbind(ecosystem %>% 
            select(time, sim, agb) %>% 
            separate(sim, c("SR", "CVH", "disturbance"), remove = F)) %>% 
    filter(SR == 25, CVH == 4, disturbance %in% c(0, 50)) %>% 
    mutate(agb = agb/10^3) %>% 
    reshape2::dcast(time ~ disturbance, value.var = "agb") %>% 
    rename(control = `0`, treatment = `50`) %>% 
    mutate(control = ifelse(is.na(control), treatment, control)) %>% 
    mutate(type = "Above-ground biomass (MgC/ha)"),
  mature %>% 
    select(time, sim, agb) %>% 
    separate(sim, c("SR", "CVH"), remove = F) %>% 
    mutate(disturbance = 50) %>% 
    rbind(ecosystem %>% 
            select(time, sim, agb) %>% 
            separate(sim, c("SR", "CVH", "disturbance"), remove = F)) %>% 
    filter(SR == 25, CVH == 4, disturbance %in% c(0, 50)) %>% 
    mutate(agb = agb/10^3) %>% 
    reshape2::dcast(time ~ disturbance, value.var = "agb") %>% 
    rename(control = `0`, treatment = `50`) %>% 
    mutate(control = ifelse(is.na(control), treatment, control)) %>% 
    mutate(treatment = treatment/control) %>% 
    mutate(control = control/control) %>%
    mutate(type = 'AGB recovery')
) %>%
  ggplot(aes(x = time)) +
  geom_line(aes(y = control, color = "control")) +
  geom_line(aes(y = treatment, color = "disturbed")) +
  geom_ribbon(aes(ymax = control, ymin = treatment), 
              fill = "lightgrey", col = NA, alpha = 0.5) +
  xlab("Time (years)") +
  ylab("") +
  scale_color_discrete("Disturbance\nintensity\n(% of BA)", 
                       labels = c("0 (control)", "50 (disturbed)")) +
  facet_wrap(~type, nrow = 2, scales = "free") +
  ggsave(filename = "./images/original2.png", width = 20, height = 20, units = "cm")
```

This document introduces and explains the recovery function $R(X)$ and the resilience index $I_R$ definitions (figure \@ref(fig:IrGraphFinal)). Resilience concept has met numerous definitions, but resilience in ecology can encompass both ecosystem ability to adsorb and recover from disturbance, resistance and recovery respectively [@Gunderson2010].

```{r IrGraphFinal, fig.cap="Definition of recovery function and resilience indice shown on aboveground biomass dynamics of a control and disturbed ecosystem before and after the disturbance event."}
knitr::include_graphics("images/figure3.png")
```

# Recovery

The recovery of the simulated forest ecosystem, i.e. the ecosystem capacity to recover from a disturbance event, at a time $t$, $R(t)$ regarding a given property $X$ was defined ad follows:

\begin{equation}
  R[X(t)] = \frac{X_D(t)}{X_C(t)}
  (\#eq:Recovery)
\end{equation}

where $X_D(t)$ and $X_C(t)$ are the ecosystem metric values at time $t$ in the disturbed simulation ($disturb_{intensity} = 25\%, 50\%, 75\%$) and in the undisturbed control simulation ($disturb_{intensity} = 0\%$).

# Resilience

When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery. We then calculated the euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$. Because the intensity of disturbance was the same for all simulations, our analysis did not include any ecosystem resistance, i.e. the ecosystem capacity to absorb a disturbance event,. Consequently **in this particular case**, we can use the cummulated recovery as an index of ecosystem resilience, further defined as $I_R$:

\begin{equation}
  I_R(X) = \int _{t=0} ^{t=600} \frac{X_D(t)}{X_C(t)}.dt
  (\#eq:Resilience)
\end{equation}

A high value of $I_R$ indicating a slow recovery. Ecosystem euclidean distance to equilibrium distance was calculated in a multi-dimensional space for the two following forest functions : (i) forest structure (AGB, BA, N, N10, and N30, e.g. figure \@ref(fig:datatrans)) and (ii) forest functioning (GPP, NPP, Rday, and Rnight).

```{r datatrans, echo=FALSE, fig.cap='Ecosystem outputs data transformation. Ecosystem outputs (**A**) are normalized by the control value over time to calculate recovery (**B**); recovery of different ecosystem outputs is then used in a multidimensional space to caluclate ecosystem distance to equilibrium (**C**); finally distance at equilibrium is integrated over time in a cumulative sum (**D**). ', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
cols <- RColorBrewer::brewer.pal(3, "YlOrRd")
g_agb <- ecosystem %>% 
  ungroup() %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb/1000, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Aboveground biomass (tonC/ha)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                       values=c("#999999", cols))
g_resilience <- resilience %>% 
  ungroup() %>%
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb, color = disturbance)) +
  geom_hline(yintercept = 1, col = 'red') +
  geom_point() +
  xlab('Time (years)') +
  ylab('Resilience of aboveground biomass') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
g_deq <- Deq %>% 
  left_join(., doe, by = 'sim') %>%
  ungroup() %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = structure, 
             color = disturbance, fill = disturbance)) +
  geom_hline(yintercept = 0, col = 'red') +
  geom_point() +
  geom_area(alpha = 0.2) +
  xlab('Time (years)') +
  ylab('Distance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols) +
  scale_fill_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
g_Ieq <- Deq %>% 
  group_by(sim) %>% 
  arrange(time) %>% 
  mutate_at(vars(-time), cumsum) %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = structure, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Cummulative integral of\ndistance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
cowplot::plot_grid(g_agb, g_resilience, g_deq, g_Ieq, labels = LETTERS[1:4])
```

# References