### 2019-01-11 S1-LL.Rmd

---
title: 'Supplementary material S1: Leaf lifespan model'
output:
  bookdown::word_document2: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: no
    toc: no
    keep_tex: yes
  bookdown::html_document2:
    number_sections: no
    toc: yes
    toc_float: yes
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r A2data, include=FALSE, echo=FALSE, cache=TRUE}
library(readr)
library(reshape2)
library(BIOMASS)
library(tidyverse)
LES <- read.csv("~/Documents/ECOFOG/TROLL/inst/extdata/LL/GLOPNET.csv", skip = 10)
LES <- LES[which(LES$BIOME == 'TROP_RF'),]
LES <- LES[which(LES$GF == 'T'),]
LES <- LES[-which(is.na(LES$log.LL)),]
data <- cbind(LES[c('Dataset', 'Species')],
              10^LES[paste0('log.',c('LL', 'LMA', 'Nmass', 'Narea',
                                     'Amass', 'Aarea', 'Gs', 'Rdmass', 'Rdarea'))])
names(data) <- c('Dataset', 'Species', 'LL', 'LMA', 'Nmass', 'Narea',
              'Amass', 'Aarea', 'Gs', 'Rdmass', 'Rdarea')
data <- droplevels(data[c('Dataset', 'Species', 'LL', 'LMA', 'Nmass')])
TRY <- read_delim("~/Documents/ECOFOG/TROLL/inst/extdata/LL/TRY3.txt",
    "\t", escape_double = FALSE, trim_ws = TRUE)
TRY <- data.frame(TRY[c('Dataset', 'SpeciesName', 'TraitName', 'StdValue')])
TRY <- TRY[-which(is.na(TRY$TraitName)),]
traits <- c('SLA', 'Nm', 'Pm', 'LL', 'wsg')
names(traits) <- unique(TRY$TraitName)
TRY$TraitName <- traits[TRY$TraitName] ; rm(traits)
K_TRY <- read_csv("~/Documents/ECOFOG/TROLL/inst/extdata/LL/Kitajima.csv")
K_TRY <- data.frame(K_TRY[c('Species', 'Leaf lifespan (months)', 'LMA (g m−2)')])
names(K_TRY) <- c('SpeciesName', 'LL', 'SLA')
K_TRY$SLA <- (1/K_TRY$SLA)*1000
K_TRY <- melt(K_TRY, id.var = c('SpeciesName'))
names(K_TRY) <- c('SpeciesName', 'TraitName', 'StdValue')
Dataset <- rep('Kitajima', dim(K_TRY)[1])
K_TRY <- cbind(Dataset, K_TRY) ; rm(Dataset)
TRY <- rbind(TRY, K_TRY) ; rm(K_TRY)
TRY <- dcast(TRY, Dataset + SpeciesName ~ TraitName,
               value.var = 'StdValue', fun.aggregate = mean, na.rm = T)
TRY <- TRY[which(!is.na(TRY$LL)),]
TRY <- TRY[c('Dataset', 'SpeciesName', 'LL', 'SLA', 'Nm', 'wsg')]
names(TRY) <- c('Dataset', 'Species', 'LL', 'LMA', 'Nm', 'wsg')
TRY$LMA <- 1/TRY$LMA*1000
TRY$Nm <- TRY$Nm/10
TRY$LMA2 <- data$LMA[match(TRY$Species, data$Species)]
TRY$Nm2 <- data$Nmass[match(TRY$Species, data$Species)]
TRY$wsg2 <- apply(do.call('rbind', strsplit(TRY$Species, ' ', fixed = T)),
                   1, function(x) getWoodDensity(x[1],
                                                 x[2], region = 'SouthAmericaTrop')$meanWD)
TRY$LMA <- apply(TRY[c('LMA', 'LMA2')], 1, mean, na.rm = T)
TRY$Nm <- apply(TRY[c('Nm', 'Nm2')], 1, mean, na.rm = T)
TRY$wsg <- apply(TRY[c('wsg', 'wsg2')], 1, mean, na.rm = T)
data <- TRY[c("Dataset", "Species", "LL", "LMA", "Nm", "wsg")] ; rm(TRY)
na <- c(which(is.na(data$LMA)), which(is.na(data$Nm)), which(is.na(data$wsg)))
data <- data[-na,]; rm(na)
LMAmax <- max(data$LMA) ; Nmax <- max(data$Nm) ; wsgmax <- max(data$wsg)
data$Dataset <- as.factor(data$Dataset)
```

# Material and methods

We compiled functional traits from GLOPNET [@wright_worldwide_2004] and DRYAD [@chave_towards_2009] databases (43 individuals, see Table \@ref(tab:A2traits)).  We measured the importance of functional traits to explain leaf lifespan with an out-of-bag method applied on a random forest. Then, we used a Bayesian approach to test different models with growing levels of complexity. We retained the model with the best trade-off between model complexity (number of parameters), convergence, likelihood, and prediction quality (root mean square error of predictions).

```{r A2traits, echo=FALSE}
data %>% 
  dplyr::select(LL, LMA, Nm, wsg) %>% 
  rename(N = Nm) %>% 
  melt(variable.name= "Abbreviation") %>% 
  group_by(Abbreviation) %>% 
  summarise(Range = paste(round(min(value),2), "-", round(max(value),2))) %>% 
  mutate(Trait = c('Leaf lifespan',
            'Leaf mass per area',
            'Leaf nitrogen',
            'Wood specific gravity')) %>% 
  mutate(Unit = c('month',' $m^2.kg^{-1}$', '$mg.g^{-1}$', '$mg.mm^{-3}$')) %>% 
  select(Trait, Abbreviation, Unit, Range) %>% 
  knitr::kable(caption = 'Gathered functional traits with abbreviation, unit, and range.', format = 'pandoc')
```

# Results

Leaf nitrogen content ($N$) was the most important trait to model leaf lifespan (see Table \@ref(tab:A2OOB)). In particular, $N$ importance was higher than leaf mass per area (158 against 96 percent of mean square error increase), which was used as a unique predictor of leaf lifespan in previous models [@Reich1991a; @Reich1997; @wright_worldwide_2004]. Finally, wood specific gravity (wsg) also improved leaf lifespan estimation.

```{r A2OOB, echo=FALSE}
randomForest::randomForest(LL ~ LMA + N + LMA + wsg,
                   data = rename(data, N = Nm), na.action = na.omit,
                   ntree = 1000, replace = F , importance = T,
                   do.trace = F, keep.forest = T, keep.inbag = T)$importance %>% 
  as.data.frame() %>% 
  rename(MSE = `%IncMSE`) %>% 
  select(MSE) %>% 
  knitr::kable(caption = 'Variable importance calculated with an out-of-bag method applied on a random forest. Variable importance is represented by the mean increase in mean square error (MSE).', format = 'pandoc', col.names = "MSE increase (%)")
```

We selected the following model:

\begin{equation}
  LL_{d} \sim log\mathcal{N}({\beta_1}_d*LMA - {\beta_2}_d*N*\beta_3*wsg, \sigma)
  (\#eq:LL)
\end{equation}

Leaf lifespan $LL$ follows a log-normal distribution. Each ${\beta_i}_d$ follows a normal distribution centered on $\beta_i$. All $\beta_i$, $\sigma_i$, and $\sigma$ are assumed without preemption following a gamma distribution. $d$ represents the dataset origin within GLOPNET database and is integrated as a random effect that encompass environmental and protocol variations.

```{r A2LLpred, echo=FALSE, fig.cap='Leaf lifespan predictions using the selected model with leaf mass per area (LMA), leaf nitrogen content (N), and wood specific gravity (wsg), and predicted versus observed values. Points represent the observed values whereas red line represent the mean predicted values and shaded areas the 5-95% confidence interval for the predicted value.', cache=TRUE, fig.height=12, fig.width=12}
library(ggplot2)
load("~/Documents/ECOFOG/TROLL/inst/doc/LL_models/M11.LL.Rdata")
fit <- fit$full
beta <- as.matrix(fit)[which.max(as.matrix(fit)[,'lp__']),c('beta_1', 'beta_2', 'beta_3')]
pred <- apply(as.matrix(fit), 1, function(x)
  sapply(seq_len(dim(data)[1]), function(i)
      rlnorm(1, x['beta_1']*(data$LMA[i]/LMAmax) - x['beta_2']*(data$Nm[i]/Nmax) + x['beta_3']*(data$wsg[i]/wsgmax), x['sigma'])
    ))
pred <- data.frame(t(apply(pred, 1, function(x) quantile(x, probs = c(0.15, 0.85)))))
names(pred) <- c('5%', '95%')
pred$mean <- exp((beta[1]*(data$LMA/LMAmax) - beta[2]*(data$Nm/Nmax) + beta[3]*(data$wsg/wsgmax)))
pred$LMA <- data$LMA
pred$Nm <- data$Nm
pred$wsg <- data$wsg
pred$obs <- data$LL
lma <- ggplot(data, aes(x = LMA, y = LL)) +
  geom_ribbon(aes(x = pred$LMA, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
  geom_line(aes(x = pred$LMA, y = pred$`5%`, color = '5%'), linetype = 2) +
  geom_line(aes(x = pred$LMA, y = pred$`95%`, color = '95%'), linetype = 2) +
  geom_line(aes(x = pred$LMA, y = pred$mean, color = 'mean'), linetype = 2) +
  geom_point() +
  theme_bw() +
  xlab('Leaf mass per area (LMA in g/m2)') +
  ylab('Leaf lifespan (LL in months)') +
  scale_color_manual(name = 'Model',
                     values = c('5%' = 'black',
                                '95%' = 'black',
                                'mean' = 'red'))
nm <- ggplot(data, aes(x = Nm, y = LL)) +
  geom_ribbon(aes(x = pred$Nm, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
  geom_line(aes(x = pred$Nm, y = pred$`5%`, color = '5%'), linetype = 2) +
  geom_line(aes(x = pred$Nm, y = pred$`95%`, color = '95%'), linetype = 2) +
  geom_line(aes(x = pred$Nm, y = pred$mean, color = 'mean'), linetype = 2) +
  geom_point() +
  theme_bw() +
  xlab('Leaf nitrogen content (Nmass in %)') +
  ylab('Leaf lifespan (LL in months)') +
  scale_color_manual(name = 'Model',
                     values = c('5%' = 'black',
                                '95%' = 'black',
                                'mean' = 'red'))
wsg<- ggplot(data, aes(x = wsg, y = LL)) +
  geom_ribbon(aes(x = pred$wsg, ymin = pred$`5%`, ymax = pred$`95%`), color = 'grey', alpha = 0.2) +
  geom_line(aes(x = pred$wsg, y = pred$`5%`, color = '5%'), linetype = 2) +
  geom_line(aes(x = pred$wsg, y = pred$`95%`, color = '95%'), linetype = 2) +
  geom_line(aes(x = pred$wsg, y = pred$mean, color = 'mean'), linetype = 2) +
  geom_point() +
  theme_bw() +
  xlab('Leaf nitrogen content (Nmass in %)') +
  ylab('Leaf lifespan (LL in months)') +
  scale_color_manual(name = 'Model',
                     values = c('5%' = 'black',
                                '95%' = 'black',
                                'mean' = 'red'))
predvsobs<- ggplot(pred, aes(x = obs, y = mean)) +
  geom_ribbon(aes(ymin =`5%`, ymax = `95%`), color = 'grey', alpha = 0.2) +
  geom_line(aes(y = `5%`, color = '5%'), linetype = 2) +
  geom_line(aes(y = `95%`, color = '95%'), linetype = 2) +
  geom_line(aes(y = mean, color = 'mean'), linetype = 2) +
  geom_point() +
  theme_bw() +
  xlab('Observed leaf lifespan (LL in months)') +
  ylab('Predicted leaf lifespan (LL in months)') +
  geom_abline(slope = 1) +
  scale_color_manual(name = 'Model',
                     values = c('5%' = 'black',
                                '95%' = 'black',
                                'mean' = 'red'))
cowplot::plot_grid(lma, nm, wsg, predvsobs, labels = c('LMA', 'Nmass', 'wsg', 'P v. O'))
```

Simulations with this new allometry have realistic estimations of above-ground biomass even with low-LMA species, resolving our initial issue. For instance with this allometry *Symphonia sp 1* (a low LMA species) is now reaching a realistic aboveground biomass above 400  $\textrm{tonC.ha}^{-1}$ and realistic diameter and age distribution inside the final population.

# References