### 2019-01-11 S3-Partitioning.Rmd

---
title: 'Supplementary material S3: ?'
output:
  bookdown::word_document2: default
  bookdown::pdf_document2:
    number_sections: false
  bookdown::html_document2:
    number_sections: false
    toc_float: yes
csl: /home/sylvain/Documents/Bibliography/csl/joe.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(kableExtra)
options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
theme_set(bayesplot::theme_default())
opts_chunk$set( echo = F, message = F, warning = F, 
                fig.height = 8, fig.width = 12,
                cache = T, cache.lazy = F)
```

```{r BE, echo=FALSE}
path <- '~/Documents/ECOFOG/Results/disturbance/disturbance/data'
load(file.path(path, 'disturbanceDOE.Rdata'))
doe <- rename(doe, SR = richness)
load(file.path(path, 'disturbanceBE.Rdata'))
load(file.path(path, 'disturbanceBErecovery.Rdata'))
```

This document explores the trajectories of biodiversity effect on aboveground biomass partitioned into complementarity and selection effects from forest establishment from a bare soil to 600 years after the disturbance event (Figure \@ref(fig:gBE)). Similarly to the aboveground biomass of the ecosystem, the net biodiversity effect and selection effect decrease following the disturbance, whereas the complementarity effect directly increases following the disturbance (sub-figure on right-bottom on figure \@ref(fig:gBE)). Interestingly, this pattern of post-disturbance decrease of selection effect, followed by a recovery phase, is similar to the effect of succession in forest early stages: when the new cohort arrives selection effect suddenly decreases and complementarity effect insures a small net effect (first century, sub-figure on left-bottom on figure \@ref(fig:gBE)). The pattern is stronger after a disturbance, maybe due to an increased diversity allowing a stronger diversity effect compared to the absence of already existing diversity when forest established from bare soil.

```{r gBE, echo=FALSE, fig.cap='Post-disturbance trajectories of complementarity and selection effects on aboveground biomass for increasing levels of disturbance (25%, 50% and 75% of basal area loss). The net biodiversity ($NE$), complementarity ($CE$) and selection ($SE$) effects were normalized by the net effect of biodiversity in the control simulation ($disturb_{intensity}=0\\%$). Solid lines represents the mean value whereas shaded areas represent the range of values across the 20 simulations with a species richness ($SR$) of 125. Time axis has been square-root transformed to highlight first decades dynamic.', fig.height=12, dpi=600}
BEresilience %>% 
  left_join(., doe, by = 'sim') %>%
  filter(SR == 125) %>% 
  select(time, BE, disturbance, SR,
         agb, ba, n, n10, n30,
         gpp, npp, Rday, Rnight) %>% 
  reshape2::melt(id.vars = c("time", "BE", "disturbance", "SR")) %>% 
  group_by(time, BE, disturbance, SR, variable) %>%
  filter(!is.na(value)) %>% 
  summarise(qt1 =quantile(value, probs=0.05),
            qt2 =quantile(value, probs=0.75),
            mean=mean(value),
            n=n()) %>% 
  ggplot(aes(x = time)) +
  geom_hline(yintercept = 1, color = 'red') +
  geom_hline(yintercept = 0, color = 'blue') +
  geom_ribbon(aes(ymin = qt1, ymax = qt2, fill = BE), alpha = 0.3) +
  geom_line(aes(y = mean, color = BE)) +
  facet_grid(variable ~ disturbance, labeller = 'label_both', scales = "free") +
  xlab('Time since disturbance (years)') +
  ylab('Biodiversity effect') +
  scale_x_sqrt() +
  theme(text = element_text(size=20),
        legend.key.size=unit(2,"line")) +
  ylim(-0.5, 1.5)
```